Bee visitation, pollination services, and plant yield in commodity and hybrid seed canola

### 2021-12-08 Manuscript/main.tex
Broke model into pieces for easier debugging

All component GLMMs of the pSEM were written in Stan 2.17.1 \citep{gelman2015}, and run using \texttt{rstan} 2.18.1.
We used informative normal priors ($\mu=0, \sigma=5$) for fixed effects, and gamma priors ($\alpha=1, \beta=1$) for the variance components.
We ran three separate chains with an adaptive phase of 2000 iterations, and a sampling phase of 2000 iterations, then checked for convergence of the chains ($\hat{R} \approx$1) and low autocorrelation within chains (high $N_{eff}$).
We assessed the adequacy of the underlying probability distribution functions of each model using posterior predictive checks \citep{gelman2013}, and found that the probability distributions were specified correctly (p$\sim$0.5) for all components except seed count.
This means that while the average seed counts (per plant) were unbiased, the variance within a plant was not. 
However, the negative binomial distribution had the closest posterior predictive checks of any other distribution. 
% p-values listed below (also in Tables S3 and S4) are \textit{posterior} p-values, representing the proportion of the posterior distribution that overlapped zero, with Z-scores (mean/standard deviation of posterior) showing the effect size.
% Predictions from models are median values of the posterior, representing the median prediction, while intervals displayed in figures are credible intervals (CIs), representing 95\% of the samples drawn from the posterior distribution (analogous to confidence intervals, see \citealt{gelman2013}).

@@

@@

The SEMs revealed that plant size and pollen deposition were the main drivers of yield in seed canola, while only plant size mattered in commodity canola.
Distance from the field edge had a negative effect on honey bee visitation in commodity and seed canola (Z=-3.7, -2.5, respectively), and unexpectedly, had a negative effect on leafcutter bee visitation (Z= -4.4), showing that overall visitation decreases with distance into the field.
The path from leafcutter bee visitation to pollen deposition was strong in seed canola (Z=3.2), but there was no path from honey bee visitation to pollen deposition in either crop type.
Pollen deposition increased fruit set (Z=3.1) and seeds per pod (Z=2.3) in seed canola, but was not present in commodity canola, showing that yield in seed canola, but not commodity canola, is strongly dependent on pollination.
%The path from seeds per pod to weight per seed was negative in seed canola (Z=-15.0) but positive in commodity canola (Z=6.1) 
Both crop types had strong negative paths from flower survival to flowers per plant (commodity Z=-12.5, seed Z=-12.1), indicating that high flower survival suppressed future flower production.
Commodity canola had a similar weight per seed, regardless of the plant size, but in seed canola, plant size had a strong influence on seeds per pod (Z=5.6) and weight per seed (4.6).
Finally, in both crop types, the paths leading from plant size were the dominant signal influencing fruit set (commodity Z=5.8; seed Z=19.9), but the path from pollen deposition to fruit set was only present in seed canola, showing that plant resources matter strongly for both crop types, but that increased pollen deposition only benefits seed canola.

\subsection{Prediction 1: Decline in visitation with distance}
Bee visitation strongly declined with distance away from the hives and shelters, in both commodity and seed canola fields. 
In commodity canola, honey bee visitation sharply declined with distance into the field (Z=-3.7, p$=$0.0002, Figure \ref{fig:hbeeDist_commodity}), decreasing from 8.5 visits/hr at the edge of the field to 3.1 visits/hr at 100 m into the field (at a stocking level of 20 hives).
Visitation rates were higher in both years in Lethbridge (2014: 9.3 visits/hr, 2015: 7.1), but were higher in 2015 than 2014 at Grande Prairie (Z=2.3, p$=$0.02), rising from 5 visits/hr at the field edge in 2014 to 17 visits/hr in 2015. % (at a stocking level of 20 hives). 
Honey bee visitation increased with the number of hives used for stocking (Z=5.3, p$<$0.0001); plot-level visitation at the edge of the field increased from 14.5 visits/hr with an apiary of 20 hives to 22.7 visits/hr with an apiary of 40 hives, while unstocked field edges had 1.8 visits/hr.
In seed canola fields, honey bee visitation also declined with distance into the field (Z=-2.5, p$=$0.01), but not as strongly as in commodity fields (126 visits/hr at edge of field vs. 78 visits/hr at 100 m). 
Honey bee visitation was lower near leafcutter shelters (Z=3.9, p$=$0.001), dropping from 60 visits/hr at 10 m to 34 visit/hr at 2 m, likely due to competition with leafcutter bees (Figure \ref{fig:hbeeDist_both}, \ref{fig:lbeeDist_both}).

Leafcutter bees were much more sensitive to distance from their shelter than honey bees.
Leafcutter bee visitation sharply declined with distance from their shelters (Z=-11.4, p$=$0.0001, 350 visit/hr at 2 m to 83 visit/hr at 10 m), and unexpectedly, was higher at the edge of the field than the centre (Z=-4.4, p$=$0.0001, 114 visits/hr vs 28 visits/hr).
However, there was no difference in visitation rates when the stocking rate of cocoons per shelter was lower (Z=2.2, p$=$0.028), indicating that leafcutters at high stocking densities may migrate to shelters close to the edge of the field (Figure \ref{fig:hbeeDist_both}).
Both leafcutter bees and honey bees visited at similar rates in the male and female bays (leafcutter bee: p$=$0.72, honey bee: p$=$0.32), but within the female bay, leafcutter bee visitation was lower in the centre of the bay (p$=$0.04, 51 vs. 29 visits/hr), while honey bee visitation was higher at the centre of the bay (p$=$0.008, 71 vs. 117 visits/hr, Figure \ref{fig:edgeCentVis_both}).
%Therefore, visitation by both species strongly decreases with distance from the edge of the field, decreases with distance from the leafcutter shelters in seed canola fields, and is done by different bees at the centre of the female bay.

@@

\subsection{Prediction 2: Pollen deposition increased by bee visitation}
Pollen deposition was enhanced by bee visitation in seed canola, but not in commodity canola.
In commodity canola, pollen deposition on stigmas was high (mean: 293 grains per stigma, SD: 385, range: 0--3981), but honey bee visitation did not affect deposition (p$=$0.77, Figure \ref{fig:hbeeVisPol}).
Pollen grains per stigma weakly declined with distance from the edge of the field (p$=$0.07), but this only amounted to an average decrease of 290 grains per stigma at the edge compared to 240 at the field centre (6\% decrease), meaning that distance did not strongly affect pollen deposition either.
Flies were common visitors in commodity fields (Table \ref{tab:propVisitors}), but had no detectable effect on pollen deposition (p$=$0.577), and other flower visitors were rare, so only honey bees and leafcutter bees were considered as potential pollinators.
In seed canola, overall pollen deposition was much lower (mean: 22 grains per stigma, SD: 43, range: 0--578), was strongly increased by leafcutter bee visitation (Z=3.2, p$=$0.002, Figure \ref{fig:visPol_both}, 12 grains/stigma at 6 visits/hr vs. 16 pollen grains at 60 visits/hr), and also decreased with distance from the edge of field (Z=-4.4, p$=$0.0001, Figure \ref{fig:distCentPol}, 23 vs. 11 grains/stigma).
Pollen deposition was lower in the centre of the female bay (Z=-4.5, p$=$0.0001, 23 vs. 13 grains/stigma) independent of visitation rate, suggesting that pollen is lost from bees crossing from male bays, who likely visit flowers at the edge of the bay first \citep{pinnisch1990}. 
There was no direct effect of honey bees on pollen deposition (Z=0.98, p$=$0.33), implying that most of the pollen deposition occurs via leafcutter visitation.

>>>>

All component GLMMs of the pSEM were written in Stan 2.17.1 \citep{gelman2015}, and run using \texttt{rstan} 2.18.1.
We used informative normal priors ($\mu=0, \sigma=5$) for fixed effects, and gamma priors ($\alpha=1.05, \beta=1$) for the variance components.
We ran four separate chains with an adaptive phase of 1000 iterations, and a sampling phase of 1000 iterations, then checked for convergence of the chains ($\hat{R} \approx$1) and low autocorrelation within chains (high $N_{eff}$).
We assessed the adequacy of the underlying probability distribution functions of each model using posterior predictive checks \citep{gelman2013}, and found that the probability distributions were specified correctly (p$\sim$0.5) for all components except seed count.
This means that while the average seed counts (per plant) were unbiased, the variance within a plant was not. 
However, the negative binomial distribution had the closest posterior predictive checks of any other distribution. 
% p-values listed below (also in Tables S3 and S4) are \textit{posterior} p-values, representing the proportion of the posterior distribution that overlapped zero, with Z-scores (mean/standard deviation of posterior) showing the effect size.
% Predictions from models are median values of the posterior, representing the median prediction, while intervals displayed in figures are credible intervals (CIs), representing 95\% of the samples drawn from the posterior distribution (analogous to confidence intervals, see \citealt{gelman2013}).

@@

Plant size and pollen deposition were the main drivers of yield in seed canola, while plant size was the main driver of yield in commodity canola.
Distance from the field edge had a negative effect on honey bee visitation in both commodity and seed canola (Z=-3.7, -2.5, respectively), and unexpectedly, had a negative effect on leafcutter bee visitation (Z= -4.4), showing that overall visitation decreases with distance into the field.
The path from leafcutter bee visitation to pollen deposition was strong in seed canola (Z=3.2), but there was no path from honey bee visitation to pollen deposition in either crop type.
Pollen deposition increased fruit set (Z=3.1) and seeds per pod (Z=2.3) in seed canola, but was not present in commodity canola, showing that yield in seed canola, but not commodity canola, is strongly dependent on pollination.
%The path from seeds per pod to weight per seed was negative in seed canola (Z=-15.0) but positive in commodity canola (Z=6.1) 
Both crop types had strong negative paths from flower survival to flowers per plant (commodity Z=-12.5, seed Z=-12.1), indicating that high flower survival suppressed future flower production.
Commodity canola had a similar weight per seed, regardless of the plant size, but in seed canola, plant size had a strong influence on seeds per pod (Z=5.6) and weight per seed (4.6).
Finally, in both crop types, the paths leading from plant size were the dominant signal influencing fruit set (commodity Z=5.8; seed Z=19.9), but the path from pollen deposition to fruit set was only present in seed canola, showing that plant resources matter strongly for both crop types, but that increased pollen deposition only benefits seed canola.

@@

Bee visitation strongly declined with distance away from the hives and shelters, in both commodity and seed canola fields. 
In commodity canola, honey bee visitation declined with distance into the field (Z=-3.7, p$=$0.0002, Figure \ref{fig:hbeeDist_commodity}), decreasing from 8.5 visits/hr at the edge of the field to 3.1 visits/hr at 100 m into the field (at a stocking level of 20 hives).
% Visitation rates were higher in both years in Lethbridge (2014: 9.3 visits/hr, 2015: 7.1), but were higher in 2015 than 2014 at Grande Prairie (Z=2.3, p$=$0.02), rising from 5 visits/hr at the field edge in 2014 to 17 visits/hr in 2015. % (at a stocking level of 20 hives). 
Honey bee visitation increased with the number of hives used for stocking (Z=5.3, p$<$0.0001); plot-level visitation at the edge of the field increased from 14.5 visits/hr with an apiary of 20 hives to 22.7 visits/hr with an apiary of 40 hives, while unstocked field edges had an average of 1.8 visits/hr.
In seed canola fields, honey bee visitation also declined with distance into the field (Z=-2.5, p$=$0.01), but not as strongly as in commodity fields (126 visits/hr at edge of field vs. 78 visits/hr at 100 m). 
Honey bee visitation was lower near leafcutter shelters (Z=3.9, p$=$0.001), dropping from 60 visits/hr at 10 m to 34 visits/hr at 2 m, likely due to competition with leafcutter bees (Figure \ref{fig:hbeeDist_both}, \ref{fig:lbeeDist_both}).

Leafcutter bees were much more sensitive to distance than honey bees, and visitation declined quickly with distance from their shelters (Z=-11.4, p$=$0.0001, 350 visit/hr at 2 m to 83 visit/hr at 10 m), and unexpectedly, was higher at the edge of the field than the centre (Z=-4.4, p$=$0.0001, 114 visits/hr vs 28 visits/hr).
However, there was no difference in visitation rates when the stocking rate of cocoons per shelter was lower (Z=2.2, p$=$0.028), indicating that leafcutters at high stocking densities may migrate to shelters close to the edge of the field (Figure \ref{fig:hbeeDist_both}).
Both leafcutter bees and honey bees visited at similar rates in the male and female bays (leafcutter bee: p$=$0.72, honey bee: p$=$0.32), but within the female bay, leafcutter bee visitation was lower in the centre of the bay (p$=$0.04, 51 vs. 29 visits/hr), while honey bee visitation was higher at the centre of the bay (p$=$0.008, 71 vs. 117 visits/hr, Figure \ref{fig:edgeCentVis_both}).
%Therefore, visitation by both species strongly decreases with distance from the edge of the field, decreases with distance from the leafcutter shelters in seed canola fields, and is done by different bees at the centre of the female bay.

@@

Leafcutter bee visitation increased pollen deposition in seed canola, but honey bees did not increase pollen deposition in either commodity or seed canola.
Pollen deposition on stigmas was high in commodity canola (mean: 293 grains per stigma, SD: 385, range: 0--3981), but honey bee visitation did not effect deposition (p$=$0.77, Figure \ref{fig:hbeeVisPol}).
Pollen deposition declined with distance from the edge of the field (p$=$0.07), but this only amounted to a 6\% average decrease (290 grains per stigma at the edge, 240 at field centre).
Flies were common visitors in commodity fields (Table \ref{tab:propVisitors}), but had no effect on pollen deposition (p$=$0.577), so only honey bees and leafcutter bees were considered as potential pollinators (other flower visitors were rare).
In seed canola, pollen deposition was much lower (mean: 22 grains per stigma, SD: 43, range: 0--578), and leafcutter bee visitation had a strong positive effect (Z=3.2, p$=$0.002, Figure \ref{fig:visPol_both}, 12 grains/stigma at 6 visits/hr vs. 16 pollen grains at 60 visits/hr), and also decreased with distance from the edge of field (Z=-4.4, p$=$0.0001, Figure \ref{fig:distCentPol}, 23 vs. 11 grains/stigma).
Pollen deposition was lower in the centre of the female bay (Z=-4.5, p$=$0.0001, 23 vs. 13 grains/stigma) independent of visitation rate, suggesting that pollen is lost from bees crossing from male bays, who likely visit flowers at the edge of the bay first \citep{pinnisch1990}. 
Finally, there was no effect of honey bees on pollen deposition in seed canola fields (Z=0.98, p$=$0.33), implying that most of the pollen deposition is the result of leafcutter visitation.
